---
title: "Return Distribution - Risk "
format: html
editor: visual
---

### Understand and interpret **Mean, Standard Deviation, Skewness, and Kurtosis**.

------------------------------------------------------------------------

## **Mean (Average Return)**

Shows the **average daily return**\
Positive = uptrend; Negative = downtrend

| Range | Interpretation              |
|-------|-----------------------------|
| \~0   | flat                        |
| \> 0  | Average gain (bullish bias) |
| \< 0  | Average loss (bearish bias) |

------------------------------------------------------------------------

## **Standard Deviation (Volatility)**

Measures how much returns vary daily\
Higher = more risk (price jumps/sinks)

| Daily Std Dev | Risk Level         |
|---------------|--------------------|
| \< 1%         | Low (stable asset) |
| 1–2%          | Moderate           |
| \> 2%         | High (tech/crypto) |

------------------------------------------------------------------------

## S**kewness (Tail Direction Risk)**

| Value | Shape      | What It Means                         | Risk Impact       |
|--------------|--------------|------------------------------|--------------|
| ≈ 0   | Symmetric  | Normal distribution                   | Neutral           |
| \> 0  | Right skew | Rare big gains, frequent small losses | Potential upside  |
| \< 0  | Left skew  | Rare big losses, frequent small gains | Watch for crashes |

> **Worry when:** Skewness \< 0\
> Suggests hidden downside risk (e.g. market crash potential)

------------------------------------------------------------------------

## **Kurtosis (Tail Size / Shock Risk)**

| Value | What It Means                         | Risk Impact              |
|-------|---------------------------------------|--------------------------|
| ≈ 3   | Normal (Gaussian) tails               | Baseline assumption      |
| \> 3  | **Fat tails** — big jumps more likely | Extreme events likely    |
| \< 3  | **Thin tails** — rare extreme returns | Safer, but may miss risk |

> **Worry when:**\
> - Kurtosis \> 10 → **extreme volatility or jumps likely**\
> - Especially dangerous **if paired with negative skewness**

------------------------------------------------------------------------

## Tips:

-   Always interpret **mean + std dev + skew + kurtosis together**
-   High kurtosis + negative skew = **“sneaky risky” asset**
-   If kurtosis is high → **use GARCH, VaR, ES, or stress tests**
-   Plot a **histogram** or **QQ plot** to visualize outliers

------------------------------------------------------------------------

### Reusable R function to describe a distribution

```{r}

describe_distribution <- function(x) {
  library(moments)
  cat("Mean:", mean(x, na.rm = TRUE), "\n")
  cat("Std Dev:", sd(x, na.rm = TRUE), "\n")
  cat("Skewness:", skewness(x, na.rm = TRUE), "\n")
  cat("Kurtosis:", kurtosis(x, na.rm = TRUE), "\n")

  skew <- skewness(x, na.rm = TRUE)
  kurt <- kurtosis(x, na.rm = TRUE)

  if (skew < 0) cat("Warning: Left skewed (risk of large negative returns)\n")
  if (kurt > 3) cat("Warning: Fat tails detected — extreme events more likely\n")
  if (kurt < 3) cat("️ Thin tails — extreme returns less likely\n")
}

```

Use it like this:

```{r}
describe_distribution(GLEN_returns$log_returns)
```

```         
```

``` r
```

### Instruction:

Use your own data set and map paths with:

#package::data set

#load(data set)

#read.csv()

#readxl()

I used file named GLEN, so easy to replace that with your own file.

Better calculate log returns and then input into moments of financial time series

GLEN %\>% mutate( simple_returns = (Price - lag(Price))/lag(Price),

log_returns = log(Price/lag(Price)) )

for GARCH, see my other models on GITHUB (<https://github.com/sahilsng007>)
